Transforming Trash as an Artistic Act

### 2016-01-17 chapter3_final.tex
Print copy preperations, proof reading
Former-commit-id: c4e7e2d53515fdd32b70a48f44d38c9e790b69b9

Firstly I look at how artist started using objects that are not produced by herself/himself and with the purpose of making art. Considered early works are notable and founding ones that open new dimensions to reconsider objects and their meaning. Further, they extended the ways of art making and the language of art. Therefore, it is important and necessary to analyze them. In other words, I first look at the foundation of using non-art objects in the art making. Early examples are analyzed and how they evolved over time is explored. I try to find out how they affected successor artist. Secondly this chapter looks some examples of contemporary works. Particularly, I try to capture the current approaches on trash. Several methods and genres are explored. Lastly focused on the inspiring documentary of Agnès Varda, The Gleaners and I, which provides a general picture of the subject.

@@

From the beginning of twentieth century-century non-art objects draw attention of artists. Usage of these objects play significant role in the development of art trough time. Usage of these objects dramatically changed the way how people look at art. In short, they introduced a new way of making art.

One of the remarkable change in twentieth century-century art is the use of unconventional materials in the making of paintings and sculptures, rather than the conventional materials produced with the purpose of art. These objects are not commonly recognized as art often because they already have a non-art function. Their place in the society and people’s mind commonly have a different meaning. Beyond that, artists started to use them for their own purposes. Thus, objects gained different meanings and usages with the hands of artists. It is explained in the previous chapter that objects are not static and absolute. Objects can be used for new purposes even if they lost their meaning and value in the conventional system.

@@

Using non-art object can be considered as moving away from the purification of art which is also significant movement in modern art. Purification argues the separation of art and the others. Abstract and minimalist works can be given as an example to the purification. However, methods such as collage and assemblage are a composition of different fragments and pieces. Rather than the traditional way of painting on canvas modern artists composed their work with the combination of various objects from different context. With this approach works can be examined in multi-dimensional context. This also represents another branch in modern art.

In twentieth century-century, modern artists experimented with new ways of seeing and with fresh ideas about the nature of materials and functions of art. Previously paintings are based on linear perspective and realistic style. Themes are often composed of religious and mythical narratives. After the second half of the 19th-century, they started to abandon these themes and painting methods. For instance, in the works of impressionist paint brushstrokes can be easily identified. In impressionist paintings, themes are composed of everyday scenes. With Cezanne, the linear perspective is started to question and later this questioning reached to its peak point with cubism. Cubism is developed with the collaboration of Picasso and Braque. Besides, these two artists experimented non-traditional materials in their works for the first time in the scope of fine art. They continuously produced works with using different objects. They support each others art. They showed that an endless potential lay there. They collaborated with each other on the development of cubism and collage. One of the keys to understand the significance of Cubism of Picasso and Braque, is to view their actions and how unusual they were for the time. When they placed mass produced artifacts into the realm of fine art they acted as artistic iconoclasts of their age.

Collage originates from the French \textit{coller} is a technique of pasting together manufactured, printed, or found materials, such as pieces of newspaper, fabric, wallpaper to a panel or canvas, frequently in combination with painting. In about 1912–13 Pablo Picasso and Georges Braque extended this technique, combining fragments of paper, oil-cloth, and newspapers with oil paint on canvas to construct compositions. Pasting paper is not a new technique but using this technique it in the art making is a revolutionary movement in the language of art \citep{waldman1992collage}. \quotes{Collage was a major turning point in the evolution of Cubism\footnote{It needs to be noted that there is two types of Cubism; analytical and synthetic. Analytical one early phase of cubism and based on fragmentary appearance of multiple viewpoints and overlapping planes. On the other hand synthetic one is the successor phase of analytical one and formed with various textures and patterns.}, and therefore a major turning point in the whole evolution of modernist art in this century} \citep{greenberg1984collage}.

Similar to the collage, assemblage is produced by the incorporation of three-dimensional objects. \cite{seitz1961art}, the curator of the exhibition \quotes{The Art of Assemblage} featured at the New York Museum of Modern Art in 1961, described assemblages as being made up of preformed natural or fabricated materials, objects, or fragments not designed as art materials.

One of the earliest and notable examples of collage is Picasso’s \textit{Still Life with Chair Caning} (Figure \ref{fig:Picasso_Chair}), in which a piece of oilcloth which is used as tablecloth or self coverings with an imitation chair caning design was pasted onto the painting. This work of Picasso also can be considered as first example of assemblage because of a rope that was used to frame the picture. Picasso combined painting with already existing materials. As mentioned in the name of the work, Picasso give clues of the parts of it. For instance the oval canvas and the rope around it represent the table. Pasted chair caning signifies the chair near the table. On the table folded paper shadow represent newspaper. Further there is a cut out lemon painted onto the table. The three letters above the scrap of cloth, \quotes{JOU}, can be understood as both the beginning of the word \quotes{JOURNAL}, referring to the customary newspaper lying across the cafe table, and as the French verb meaning \quotes{to play.} The new technique of collage provided new possibilities of playfulness. This was a new way of making art; instead of painting an object or landscape, you could paste whatever it was right onto the canvas.

\begin{figure}[h!]
  \centering
  \includegraphics[height=6cm]{graphics/picasso_chair.png}
  \caption{Pablo Picasso, \textit{Still Life with Chair Caning}, 1912, oil on oil-cloth over canvas edged with rope, 29 x 37 cm (Musée Picasso)}
  \label{fig:Picasso_Chair}
\end{figure}

Moreover, Braque and Picasso questioned the elitism of the art world, which had always dictated the separation of common, everyday experience from the contemplative realm of artistic creation. Of equal importance, their work highlighted and separated the role of technical skill from art-making. Braque and Picasso introduced a fake element on purpose, not to mislead or fool their audience, but rather to force a discussion of art and craft, of unique and mass-produced objects. They inquiry as \quotes{Can this object still be art if I don’t actually render its forms myself, if the quality of the art is no longer directly tied to my technical skills or level of craftsmanship?}

After Picasso and Baraque, many artists applied collage and assemblage techniques to their works. Dada artist Kurt Schwitters is one of the productive artists among them. In 1918 he created his own form of Dada in Germany named \textit{Merz}, using rubbish elements such as newspapers, bus tickets and pieces of broken wood in his collages and constructions. The nonsense word \textit{Merz} is generated from the name of a bank: the Kommerz und Privatbank. \textit{Merz} was soon to become a kind of brand name for almost all his productions. \cite{webster2011kurt} puts significance of Schwitters in the development of art as follows:

\begin{quote}
The language of \textit{Merz} now finds common acceptance and today there is scarcely an artist working with materials other than paint who does not refer to Schwitters in some way. In his bold and wide-ranging experiments he can be seen as the grandfather of Pop Art, Happenings, Concept Art, Fluxus, multimedia art and post-modernism...

Schwitters’ work inspired such post-war pioneers as Jasper Johns, Robert Rauschenberg and Joseph Beuys, and he is now seen as the grandfather of post-1945 art movements, from Pop Art to Fluxus, Conceptual Art to site-specific art, and the forerunner of present day artists such as Thomas Hirschhorn, Gregor Schneider and Rachel Whiteread.
\end{quote}

@@

\begin{quote}
The German avant-garde was working from ruins literally and metaphorically, and trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. Schwitters is able to use the ruined, the waste products, as an anthropological exploration of society from both its unpleasant outcomes and its decay. \ldots trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. As he wrote: \singlequotes{It grows more or less according to the principle of a metropolis.} The \textit{Merzbau} was itself a city; and just as Marx wrote that it was not the materiality of the object but the social relations that create value, the use of urban detritus in particular, the squalid results of mass-produced human relations, infuses the materiality of Schwitters’ work with an anthropological quality. Material has transformed into information, and \singlequotes{how} has surpassed \singlequotes{what} we see. The grottos in the \textit{Merzbau} that still reveal this detritus most clearly could not be re-created in Bissegger’s reconstruction because, arguably, they are an exploration of absence, an exploration of ruin.
\end{quote}

@@

Collages and assemblages can be composed of all types of objects such as trash and found object. The notion of found object is researched by many scholars \cite{camic2010trashed, gascoyne1935short}, but usage in the art has gained new dimensions with the Duchamp’s concept of ready-made. Industrially manufactured or anything found  anymore can found a place in the art. The most notable example is \textit{Fountain} (discussed in the previous chapter), a standard urinal purchased from a hardware store and displayed on a pedestal, resting on its side. Before that first one is \textit{Bicycle Wheel} (Figure \ref{fig:Duchamp_BicycleWheel}). He wrote that the idea of a bicycle wheel upside down onto a stool. Although he claimed to select objects for his ready-made without regard to beauty, he said, \quotes{To see that wheel turning was very soothing, very comforting\ldots I enjoyed looking at it, just as I enjoy looking at the flames dancing in a fireplace.} These samples were the beginning of 
shift from an art that is striving for beauty to a higher complex or hidden meaning beyond what was seen from common artifacts found in everyday living.

@@

In the French film Amélie\footnote{Original title: Le fabuleux destin d’Amélie Poulain, in english: The Fabulous Destiny of Amélie Poulain} directed by Jean-Pierre Jeunet in 2001, young male character collects photos that are torn up and thrown away by their owner near photo booths spread through train and metro stations of Paris. He collects them and glue them together in a photo album. In the film photo booths and found photos are centrally catalytic. Amélie falls in love with him when she realize him searching for found photos. After his photo album passed to Amélie, she looks every page of photo album again and again. Through the images of unknown people, she establishes narratives that point out the significance of these images even if they are thrown away. The important part of the film, images of a man are generated who scavenge the photos of another people. There are a lot of artist who collects discarded items but their process is not well documented. Therefore we are not fully aware of their artistic process.

\begin{figure}[h!]
  \centering
  \includegraphics[height=6cm]{graphics/amelie_01.jpg}
  \caption{Still image from Amélie by Jean-Pierre Jeunet}
  \label{fig:Amelie}
\end{figure}

Prior to film Amélie, British artist Dick Jewell published a book in title of \quotes{Found Photos} in 1977. It is a collection of photo booth images that had been thrown away or torn up by the people in the photos. \cite{walker2010dick} comments on that it is \quotes{always a joy to look through.} Images capture diverse range of activities and expressions of people using photo booth machine. They are scavenged by the artist from the photo booths of London. The reasons of why photos are rejected vary including technical problems like flash or posing time. On the other hand, some of them are very good condition, and it is a mystery that why they are rejected. Walker explains that the:

\begin{quote}
gap between our exterior perception of the image and what must have been the internal dissatisfaction of the original subject results in some of the effective moments in Found Photos, where the images are not only bizarre and wonderfully funny (as many of them are) but also profoundly moving.
\end{quote}

Walker argues that even if photo booths are mechanic and simple application of photography while capturing everyday life, leave behind such ambiguous and mysteries that some people trace them. \quotes{Dick Jewell’s approach with Found Photos was less interventionist, presenting the images themselves in the spirit of either an anthropologist showing his finds or a conceptual artist claiming bits of the real world as ready-mades.} His mode of discovery includes climbing to look at the top of the photo booths and looking for the nearest waste bin. Badger describes Found Photos as \quotes{both a conceptual artwork and a sociological footnote.}

@@

On the other side for another artist photographs serve a function to record his daily trash. Tim Gaudreau photographed every object that he threw away for one year. At the end around five thousand images are covered all the gallery and opened to visitors. He notes that it just started as a documentation of his discards, but later it became (part of his life and) an obsession that he can not able to easily quit off. He says that \quotes{this collection of images intimately displays what I do, what I consume. It reveals me.} Like in the case of Tracey Emin trash turns to a medium that is used to express oneself. People are not only composed of the listed items in the CVs, economical power like pointed out in the book (and also film) Fight Club Tyler Durden draws attention to the narrator by saying: \quotes{You are not your bank account.} It is an effective way of looking things left behind rather that things put forward. Through this project, artist reached better understanding what he is and notes that this project changed his consumption choices and patterns.

@@

Filomena Cruz turns her lens to the city and its trashes that are left behind, thrown away and squeezed on the sidewalk or pavement. In her photographic series \textit{Road Kill}, she captures tiny trash corpses that are generally ignored and out of sight. In other words, the skin of city revealed through her photos in micro scale. On the skin of city there are lots of trash including \quotes{a piece of chewing gum with an engraved leaf; a flattened-out tube; a corroding paper napkin with a still intact heart; or a frog-green Crayola melting in the heat, all speak the language of \singlequotes{worthlessness} suddenly becoming meaningful} as looked through the window of they capture the city life and skin of the city that we touched. She shows us tiny details of huge cities. Somehow these trashes for all effort succeeded to stay in the towns. These trashes can not be removed from the city but moved ones are stacked on landfill, and they are not tiny as they are.

@@

In the documentary film, Waste Land, Vik Muniz frames portraits of \textit{catadores} who are pickers of recyclable materials on the dump. He visits them at one of the biggest open-air garbage dump outside of the Rio de Janeiro home of the millions of people. Film documents whole process of development and production of \textit{Pictures of Garbage} made of collected trash from dump. He selects unusual materials and people for portraits rather than the powerful and notable people as commonly encountered at Renaissance. To create portraits of \textit{catadores} there is any other material than trash because of they are already build up their life in trash of city. Also these images are sold an auction and earned money donated to the \textit{catadores}. The images of trash moved to beloved places such as homes and museums. As stated in the rubbish theory they transformed to a durable object.

\begin{figure}[h!]
  \centering
  \includegraphics[height=10cm]{graphics/vik_muniz_marat.jpg}
  \caption{Vic Muniz, \textit{Marat}, 2008, mixed media, }
  \label{fig:VicMuniz_PicturesOfGarbage}
\end{figure}

Not only monumental portraits are extracted from the dump but also an orchestra is grounded up from the dump of Paraguay pioneered by Favio Chávez and Nicolás Gómez. Favio Chávez is a music teacher, and Nicolás Gómez is a garbage picker. They live in Cateura at Paraguay. It is a one of the poorest villages where people live among the sea of garbage. \quotes{A violin is worth more than a house here,} says Favio Chávez by pointing out their hard conditions. Favio Chávez director and founder of the orchestra whose instruments such as violins, flutes and cellos are built from collected trash by Nicolás Gómez. Players of the orchestra are compiled from the children who are hopeless for their future. In the middle of such an existence, these people and musicians have created something both special and truly awe-inspiring. They are given birth to an orchestra that plays the masterpieces of classical music composed by Mozart, Beethoven, and Vivaldi. Trash turned to treasure and totally ironic that a signifier of high culture like classical music can be even played with imperfect instruments. For them and also, most of the people it is not affordable to buy an instrument and get the education of classical music. What they create is actually not expected from them, and their success is not limited with their location. They have given concerts around the world and their motto is: \quotes{The world send us garbage. We send back music.}

\begin{figure}
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{graphics/landfill_harmonic-sax.jpg}
    \caption{Sax}
    \label{fig:landfill_harmonic-sax}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{graphics/landfill_harmonic-violin.jpg}
    \caption{Violin}
    \label{fig:landfill_harmonic-violin}
  \end{subfigure}
  \caption{Instruments of Recycled Orchestra}\label{fig:animals}
\end{figure}

@@

In the scope of painting different than the works in the previous chapter (paintings on tea bags and coffee cups) historical portraits of notable people are painted onto the flattened aluminum drink cans which are disposable objects. The people in the figure \ref{fig:Alsbrooks} is General Jefferson C. Davis who was an officer at United States Army in 19th-century. Selected persons can be viewed as a member of the higher or ruling class of society. Therefore Kim Alsbrooks, painter of these works, might have thought that to make trash is more valuable, their face will be good or vice verse. She probably use trash to criticize social elite of that time.

\begin{figure}[h!]
  \centering
  \includegraphics[height=8cm]{graphics/Alsbrooks.jpg}
  \caption{Kim Alsbrooks, \textit{Gen. Jefferson Davis}, 2015, oil on aluminum can}
  \label{fig:Alsbrooks}
\end{figure}

@@

The film American Beauty, which features a long, poetic clip of a plastic bag swirling on an eddy of air yet people find it hard to think of plastic bags as things of beauty. However, as a product --as something created and then unleashed to become seamlessly integrated into the lives of millions of people around the world--- there is a strange allure to them, just as a pathologist can admire the structure of a particularly virulent and contagious virus.

% TODO burayı da düzeltmek gerek.
Examples are not limited with these. Even if \quotes{trash is dirty and smelly trash can provide the raw materials for exquisite art---from sculpture to film and beyond.}

%
\section{The Documentary \quotes{The Gleaners and I} by Agnès Varda}
\quotes{The Gleaners and I} is a French documentary film directed by Agnès Varda in 2000. The film emphasizes diverse types of gleaning from rural to urban. According to the Merriam-Webster dictionary gleaning refers \quotes{to gather grain or other produce left by reapers}. Varda searches for the historical and current practices of gleaning in rural and urban areas of France. Historically, gleaning is the act of collecting what is left and not profitable crops after the harvest. By combining different fragments through his film, she shows that the act of gleaning also exists in various forms and styles. In particularly she curiously focuses on the (un)familiar people who glean leftover of society. Moreover, it is a self-reflexive film because the director establishes a relationship with the practice of gleaners and her film making practice. Some people glean crops; the others discarded food, and Varda gleans images.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[height=6cm]{graphics/AgnesVarda_Potatoes.jpg}
  \caption{Still image from The Gleaners and I}
  \label{fig:AgnesVarda_Potatoes}
\end{figure}

>>>>

Firstly I look at how artist started using objects that are not produced by herself/himself and with the purpose of making art. Considered early works are notable and founding ones that open new dimensions to reconsider objects and their meaning. Further, they extended the ways of art making and the language of art. Therefore, it is important and necessary to analyze them. In other words, I first look at the foundation of using non-art objects in the art making. Early examples are analyzed and how they evolved over time is explored. I try to find out how they affected successor artist. Secondly this chapter looks some examples of contemporary works. Particularly, I try to capture the current approaches on trash. Several methods and genres are explored. Lastly focused on the inspiring documentary of Agnès Varda, \textit{The Gleaners and I}, which provides a general picture of the subject.

@@

From the beginning of twentieth century non-art objects draw attention of artists. Usage of these objects play significant role in the development of art trough time. Usage of these objects dramatically changed the way how people look at art. In short, they introduced a new way of making art.

One of the remarkable change in twentieth century art is the use of unconventional materials in the making of paintings and sculptures, rather than the conventional materials produced with the purpose of art. These objects are not commonly recognized as art or art material often because they already have a non-art function and place in society. Their place in the society and people’s mind commonly have a different meaning. Beyond that, artists started to use them for their own purposes. Thus, objects gained different meanings and usages with the hands of artists. It is explained in the previous chapter that objects are not static and absolute. Objects can be used for new purposes even if they lost their meaning and value in the conventional system.

@@

Using non-art object can be considered as moving away from the purification of art which is also significant movement in modern art. Purification argues the separation of art and the others. Abstract and minimalist works can be given as an example to the purification. However, methods such as collage and assemblage are a composition of different fragments and pieces. Rather than the traditional way of painting on canvas modern artists composed their work with the combination of various objects from different context. With this approach works can be examined in multi-dimensional context. This also represents another key branch in modern art.

Artists, in twentieth century, experimented with new ways of seeing and with unusual ideas about the functions of materials and nature of art. Previously paintings are based on linear perspective and realistic style. Themes are often composed of religious and mythical narratives. After the second half of the nineteenth century, they started to abandon these themes and painting methods. For instance, in the works of impressionist paint brushstrokes can be easily identified. In impressionist paintings, themes are composed of everyday scenes. With Cezanne, the linear perspective is started to question and later this questioning reached to its peak point with cubism. Cubism is developed with the collaboration of Picasso and Braque. Besides, these two artists experimented non-traditional materials in their works for the first time in the scope of fine art. They continuously produced works with using different objects. They support each others art. They showed that an endless potential lay there. They collaborated with each other on the development of cubism and collage. One of the keys to understand the significance of Cubism of Picasso and Braque, is to view their methods as unconventional for their time. When they placed mass produced artifacts into the realm of fine art they acted as artistic iconoclasts of their age.

Collage originates from the French \textit{coller} is a technique of pasting together manufactured, printed, or found materials, such as pieces of newspaper, fabric, wallpaper to a panel or canvas, usually in combination with drawing and painting. In about 1912–13 Pablo Picasso and Georges Braque applied this technique, combining fragments of paper, oil-cloth, and newspapers with oil paint on canvas to construct compositions. Pasting paper is not a new technique but using this technique it in the art making is a revolutionary movement in the language of art \citep{waldman1992collage}. \quotes{Collage was a major turning point in the evolution of Cubism\footnote{It needs to be noted that there is two types of Cubism; analytical and synthetic. Analytical one early phase of cubism and based on fragmentary appearance of multiple viewpoints and overlapping planes. On the other hand synthetic one is the successor phase of analytical one and formed with various textures and patterns.}, and therefore a major turning point in the whole evolution of modernist art in this century} \citep{greenberg1984collage}.

Similar to the collage, assemblage is produced by the incorporation of three-dimensional objects. \cite{seitz1961art}, the curator of the exhibition \quotes{The Art of Assemblage} featured at the New York Museum of Modern Art in 1961, describes assemblages as \quotes{being made up of preformed natural or [fabricated] materials, objects, or fragments not designed as art materials.}

One of the earliest and notable examples of collage is Picasso’s \textit{Still Life with Chair Caning} (Figure \ref{fig:Picasso_Chair}), in which a piece of oilcloth which is used as tablecloth or self coverings with an imitation chair caning design was pasted onto the painting. This work of Picasso also can be considered as first example of assemblage because of a rope that was used to frame the picture. Picasso combined painting with already existing materials. As mentioned in the name of the work, Picasso give clues of the parts of it. For instance the oval canvas and the rope around it represent the table. Pasted chair caning signifies the chair near the table. On the table folded paper shadow represent newspaper. Further there is a cut out lemon painted onto the table. The three letters above the scrap of cloth, \quotes{JOU}, can be understood as both the beginning of the word \quotes{JOURNAL}, referring to the customary newspaper lying across the cafe table, and as the French verb meaning \quotes{to play.} In essence, the new method; collage provides new possibilities of playfulness to the artists. This was a new course of making art; instead of painting an object or landscape, you could paste whatever it was right onto the canvas.

\begin{figure}[h!]
  \centering
  \includegraphics[height=6cm]{graphics/picasso_chair.png}
  \caption{Pablo Picasso, \textit{Still Life with Chair Caning}, 1912, Oil on oil-cloth over canvas edged with rope, 29 x 37 cm (Musée Picasso)}
  \label{fig:Picasso_Chair}
\end{figure}

Moreover, Braque and Picasso challenged the elitism of the art world, which had constantly forced the separation of ordinary, everyday experience from the sophisticated world of artistic creation. At the same time, their work underlined and separated the role of technical skill from art-making. Braque and Picasso introduced usage of unoriginal element on purpose, not to mislead or fool their audience, but rather to force a discussion of art and craft, of unique and mass-produced objects. They inquiried as \quotes{Can this object still be art if I don’t actually render its forms myself, if the quality of the art is no longer directly tied to my technical skills or level of craftsmanship?}

After Picasso and Baraque, many artists applied collage and assemblage techniques to their works. Dada artist Kurt Schwitters is one of the productive artists among them. In 1918 he created his own form of Dada in Germany named \textit{Merz}, using rubbish elements such as newspapers, bus tickets and pieces of broken wood in his collages and constructions. The nonsense word \textit{Merz} is generated from the name of a bank: the Kommerz und Privatbank. \textit{Merz} was soon to become a kind of brand name for almost all his productions. \cite{webster2011kurt} puts significance of Schwitters in the development of art as \quotes{The language of \textit{Merz} now finds common acceptance and today there is scarcely an artist working with materials other than paint who does not refer to Schwitters in some way.} He notes that \quotes{bold and wide-ranging experiments} of Schwitters \quotes{inspired such post-war pioneers as Jasper Johns, Robert Rauschenberg and Joseph Beuys}. Schwitters can be seen as \quotes{can be seen as the grandfather of Pop Art, Happenings, Concept Art, Fluxus, multimedia art and post-modernism.}

@@

\begin{quote}
The German avant-garde was working from ruins literally and metaphorically, and trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. Schwitters is able to use the ruined, the waste products, as an anthropological exploration of society from both its unpleasant outcomes and its decay. \ldots trash was both practically and freely available; to use it was an action that took the ruins of our society, its discarded, to question how meaning is constructed. As he wrote: \quotes{It grows more or less according to the principle of a metropolis.} The \textit{Merzbau} was itself a city; and just as Marx wrote that it was not the materiality of the object but the social relations that create value, the use of urban detritus in particular, the squalid results of mass-produced human relations, infuses the materiality of Schwitters’ work with an anthropological quality. Material has transformed into information, and \quotes{how} has surpassed \quotes{what} we see. The grottos in the \textit{Merzbau} that still reveal this detritus most clearly could not be re-created in Bissegger’s reconstruction because, arguably, they are an exploration of absence, an exploration of ruin.
\end{quote}

@@

Collages and assemblages can be composed of all types of objects such as trash and found object. The notion of found object is researched by many scholars \cite{camic2010trashed, gascoyne1936short}, but usage in the art has gained new dimensions with the Duchamp’s concept of ready-made. Industrially manufactured or anything found  anymore can found a place in the art. The most notable example is \textit{Fountain} (discussed in the previous chapter), a typical porcelain urinal purchased from a hardware store and displayed on a pedestal, lying on its side. Before that first one is \textit{Bicycle Wheel} (Figure \ref{fig:Duchamp_BicycleWheel}). He wrote that the idea of a bicycle wheel upside down onto a stool. Even though he stated to select objects for his ready-made without paying attention to beauty, he wrote, \quotes{To see that wheel turning was very soothing, very comforting\ldots I enjoyed looking at it, just as I enjoy looking at the flames dancing in a fireplace.} These samples were the beginning of shift from an art that is striving for beauty to a higher complex or hidden meaning beyond what was seen from common artifacts found in everyday living.

@@

In the French film \textit{Amélie}\footnote{Original title: Le fabuleux destin d’Amélie Poulain, in english: The Fabulous Destiny of Amélie Poulain} directed by Jean-Pierre Jeunet in 2001, young male character collects photos that are torn up and thrown away by their owner near photo booths spread through train and metro stations of Paris. He collects them and glue them together in a photo album. In the film photo booths and found photos are centrally catalytic. Amélie falls in love with him when she realize him searching for found photos. After his photo album passed to Amélie, she looks every page of photo album again and again. Through the images of unknown people, she establishes narratives that point out the significance of these images even if they are thrown away. The important part of the film, images of a man are generated who scavenge the photos of another people. There are a lot of artist who collects discarded items but their process is not well documented. Therefore we are not fully aware of their artistic process.

\begin{figure}[h!]
  \centering
  \includegraphics[height=6cm]{graphics/amelie_01.jpg}
  \caption{Still image from \textit{Amélie} by Jean-Pierre Jeunet}
  \label{fig:Amelie}
\end{figure}

Prior to film \textit{Amélie}, British artist Dick Jewell published a book in title of \textit{Found Photos} in 1977. It is a collection of photo booth images that had been thrown away or torn up by owner of the photos. \textit{Found Photos} is described by  \citet[222]{badger2014photobook} as \quotes{both a conceptual artwork and a sociological footnote.} \cite{walker2010dick} comments on that it is \quotes{always a joy to look through.} Images capture diverse range of activities and expressions of people using photo booth machine. They are scavenged by the artist from the photo booths of London. The reasons of why photos are rejected vary including technical problems like flash or posing time. On the other hand, some of them are very good condition, and it is a mystery that why they are rejected. Walker points out that the:

\begin{quote}
gap between our exterior perception of the image and what must have been the internal dissatisfaction of the original subject results in some of the effective moments in \textit{Found Photos}, where the images are not only bizarre and wonderfully funny (as many of them are) but also profoundly moving.
\end{quote}

Walker argues that even if photo booths are mechanic and simple application of photography while capturing everyday life, leave behind such ambiguous and mysteries that some people trace them. \quotes{Dick Jewell’s approach with Found Photos was less interventionist, presenting the images themselves in the spirit of either an anthropologist showing his finds or a conceptual artist claiming bits of the real world as ready-mades.} His mode of discovery includes climbing to look at the top of the photo booths and looking for the nearest waste bin.

@@

On the other side for another artist photographs serve a function to record his daily trash. Tim Gaudreau photographed every object (Figure \ref{fig:TimGaudreau_SelfPotraitRevealedByTrash}) that he threw away for one year. At the end around five thousand images are covered all the gallery and opened to visitors. He notes that it just started as a documentation of his discards, but later it became (part of his life and) an obsession that he can not able to easily quit off. He says that \quotes{this collection of images intimately displays what I do, what I consume. It reveals me.} Like in the case of Tracey Emin trash turns to a medium that is used to express oneself. People are not only composed of the listed items in the CVs, economical power like pointed out in the book (and also film) \textit{Fight Club} Tyler Durden draws attention to the narrator by saying: \quotes{You are not your bank account.} It is an effective way of looking things left behind rather that things put forward. Through this project, artist reached better understanding what he is and notes that this project changed his consumption choices and patterns.

@@

Filomena Cruz turns her lens to the city and its trashes that are left behind, thrown away and squeezed on the sidewalk or pavement. In her photographic series \textit{Road Kill} (Figure \ref{fig:FilomenaCruz_RoadKill_ReVista}), she captures tiny trash corpses that are generally ignored and out of sight. In other words, the skin of city revealed through her photos in micro scale. On the skin of city there are lots of trash including \quotes{a piece of chewing gum with an engraved leaf; a flattened-out tube; a corroding paper napkin with a still intact heart; or a frog-green Crayola melting in the heat, all speak the language of \singlequotes{worthlessness} suddenly becoming meaningful} as looked through the window of they capture the city life and skin of the city that we touched. She shows us tiny details of huge cities. Somehow these trashes for all effort succeeded to stay in the towns. These trashes can not be removed from the city but moved ones are stacked on landfill, and they are not tiny as they are.

@@

In the documentary film, \textit{Waste Land}, Vik Muniz frames portraits of \textit{catadores} who are pickers of recyclable materials on the dump. He visits them at one of the biggest open-air garbage dump outside of the Rio de Janeiro home of the millions of people. Film documents whole process of development and production of \textit{Pictures of Garbage} made of collected trash from dump. He selects unusual materials and people for portraits rather than the powerful and notable people as commonly encountered at Renaissance. To create portraits of \textit{catadores} there is any other material than trash because of they are already build up their life in trash of city. Also these images are sold an auction and earned money donated to the \textit{catadores}. The images of trash moved to beloved places such as homes and museums. As stated in the rubbish theory they transformed to a durable object.

Figure \ref{fig:VicMuniz_PicturesOfGarbage} is one of the works of Muniz. It built up piling trash together in a very large room and later photograph of it taken. It is a reproduction of \textit{The Death of Marat} (1793, Oil on canvas, 165 cm x 128 cm) by Jacques-Louis David. Painting represents journalist lying dead in his bath and it is one of the most famous images of the French Revolution.

\begin{figure}[h!]
  \centering
  \includegraphics[height=10cm]{graphics/vik_muniz_marat.jpg}
  \caption{Vik Muniz, \textit{Marat (Sebastiao), Pictures of Garbage}, 2008, Digital C print}
  \label{fig:VicMuniz_PicturesOfGarbage}
\end{figure}

Not only monumental portraits are extracted from the dump but also an orchestra is grounded up from the dump of Paraguay pioneered by Favio Chávez and Nicolás Gómez. Favio Chávez is a music teacher, and Nicolás Gómez is a garbage picker. They live in Cateura at Paraguay. It is a one of the poorest villages where people live among the sea of garbage. \quotes{A violin is worth more than a house here,} says Favio Chávez by pointing out their hard conditions. Favio Chávez director and founder of the orchestra whose instruments such as violins (Figure \ref{fig:landfill_harmonic}), flutes and cellos are built from collected trash by Nicolás Gómez. Players of the orchestra are compiled from the children who are hopeless for their future. In the middle of such an existence, these people and musicians have created something both special and truly awe-inspiring. They are given birth to an orchestra that plays the masterpieces of classical music composed by Mozart, Beethoven, and Vivaldi. Trash turned to treasure and totally ironic that a signifier of high culture like classical music can be even played with imperfect instruments. For them and also, most of the people it is not affordable to buy an instrument and get the education of classical music. What they create is actually not expected from them, and their success is not limited with their location. They have given concerts around the world and their motto is: \quotes{The world send us garbage. We send back music.}

\begin{figure}
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{graphics/landfill_harmonic-sax.jpg}
    %\caption{Sax}
    \label{fig:landfill_harmonic-sax}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{graphics/landfill_harmonic-violin.jpg}
    %\caption{Violin}
    \label{fig:landfill_harmonic-violin}
  \end{subfigure}
  \caption{Instruments of Recycled Orchestra}
  \label{fig:landfill_harmonic}
\end{figure}

@@

In the scope of painting different than the works in the previous chapter (paintings on tea bags and coffee cups) historical portraits of notable people are painted onto the flattened aluminum drink cans which are disposable objects. The people in the figure \ref{fig:Alsbrooks} is General Jefferson C. Davis who was an officer at United States Army in nineteenth century. Selected persons can be viewed as a member of the higher or ruling class of society. Therefore Kim Alsbrooks, painter of these works, might have thought that to make trash is more valuable, their face will be good or vice verse. She probably use trash to criticize social elite of that time.

\begin{figure}[h!]
  \centering
  \includegraphics[height=8cm]{graphics/Alsbrooks.jpg}
  \caption{Kim Alsbrooks, \textit{Gen. Jefferson Davis}, 2015, Oil on aluminum can}
  \label{fig:Alsbrooks}
\end{figure}

@@

Examples are not limited with these. Even if \quotes{trash is dirty and smelly trash can provide the raw materials for exquisite art---from sculpture to film and beyond.} The film American Beauty, which features a long, poetic clip of a plastic bag swirling on an eddy of air yet people find it hard to think of plastic bags as things of beauty. On other side, Eloísa Cartonera a cooperative in Buenos Aires, proudly produces handmade books of Latin American literature with cardboard covers picked from the streets. Moreover, Turkish artist İpek Duben works with found objects and has a series of head sculpture. Examples can be extended but this thesis could not cover all of them.

%****************************************
\section{The Documentary \quotes{The Gleaners and I} by Agnès Varda}
\quotes{The Gleaners and I} is a French documentary film directed by Agnès Varda in 2000. The film emphasizes diverse types of gleaning from rural to urban. According to the Merriam-Webster dictionary gleaning refers \quotes{to gather grain or other produce left by reapers}. Varda searches for the historical and current practices of gleaning in rural and urban areas of France. Historically, gleaning is the act of collecting what is left and not profitable crops after the harvest. By combining different fragments through his film, she shows that the act of gleaning also exists in various forms and styles. In particularly she curiously focuses on the (un)familiar people who glean leftover of society. Moreover, it is a self-reflexive film because the director establishes a relationship with the practice of gleaners and her film making practice. Some people glean crops; the others discarded food, and Varda gleans images.

@@

\begin{figure}[h!]
  \centering
  \includegraphics[height=6cm]{graphics/AgnesVarda_Potatoes.jpg}
  \caption{Still image from \textit{The Gleaners and I} by Agnès Varda}
  \label{fig:AgnesVarda_Potatoes}
\end{figure}
